---
title: "price_prediction"
author: "Travis Barton"
date: "5/27/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(prophet)
library(readr)
library(dygraphs)
library(xts)
library(lubridate)

library(lemon)
library(data.table)
knit_print.data.frame <- lemon_print
```

# Collect the data and estimated quanities

The following are from the bureau of labor statistics, but your personal records are a more reliable as they actually list the materials you use. 
```{r load data, results='hide', message=FALSE, render=lemon_print}
INPUT_DATE = as.POSIXct('2021-07-31', tz = 'GMT')  # make sure the format is Year-month-day


shingle_price <- read_csv("asphalt_shingle_and_coating_materials.csv")
lumber_price <- read_csv('lumber.csv')
lumber_price = lumber_price[lumber_price$DATE > as.Date('2000-01-01'),]  # only recent data
shingle_price = shingle_price[shingle_price$DATE > as.Date('2000-01-01'),]  # same

colnames(shingle_price) = c('ds', 'y')
colnames(lumber_price) = c('ds', 'y')
tail(shingle_price, 10)

```

```{r details, echo=FALSE}

find_price = function(forecast, amount, CI, date_needed){
  date_needed = as.POSIXct(date_needed)
  date_low = date_needed - days(30)
  date_high = date_needed + days(30)
  forecast = forecast[(forecast$ds > date_low) & (forecast$ds < date_high),]
  cons_forecast_price_high = max(forecast$yhat_upper)
  cons_forecast_price_low = min(forecast$yhat_lower)
  forecast_price_high = max(forecast$yhat)
  forecase_price_low = min(forecast$yhat)
  
  ret = data.frame(matrix(c(forecase_price_low*amount, forecast_price_high*amount, cons_forecast_price_low*amount, cons_forecast_price_high*amount), ncol=2, byrow = TRUE))
  colnames(ret) = c('low est.', 'high est.')
  rownames(ret) = c('casual', 'confident')
  return(ret)
}

```
# Fit the model and find the bounds 

The trends of cost are a time series analysis problem similar to predicting stocks and futures, meaning they are reliable as long as __new__ outside factors do not play a role in their prices. Take a look at lumber for ex.

```{r prophet_model, echo=FALSE, message=FALSE, warning=FALSE}

m_shingle = prophet(shingle_price)

future_shingle = make_future_dataframe(m_shingle, periods = 365, freq = "D")

forecast_shingle = predict(m_shingle, future_shingle)


m_lumber = prophet(lumber_price)

future_lumber = make_future_dataframe(m_lumber, periods = 365, freq = 'D')

forecast_lumber = predict(m_lumber, future_lumber)
```


```{r plot componants and hist lumber, echo=FALSE, warning=FALSE}

prophet_plot_components(m_lumber, forecast_lumber)
dyplot.prophet(m_lumber, forecast_lumber)


```


Because the shortages were unexpected, the model does not think the price will stay high for long, and assumes they are apart of the regular business cycle. 


```{r range estimate, echo=FALSE}
lumber_requirement = 100  # these are unitless in my mind, so lets go with 100 'lumbers'
shingle_reauirement = 1000  # similar with shingles
project_date = '2021-06-15'

prices_lumber = find_price(forecast_lumber, lumber_requirement, date_needed = project_date)


prices_shingle = find_price(forecast_shingle, shingle_reauirement, date_needed = project_date)

```

### lumber est for 100 'lumbers'

```{r echo=FALSE, render=lemon_print, echo=FALSE}
head(prices_lumber, 2)
```


```{r plot componants and hist lumber actual, echo=FALSE, warning=FALSE}

prophet_plot_components(m_lumber, forecast_lumber)
dyplot.prophet(m_lumber, forecast_lumber)


```



```{r price_estimate_hist_shingle, echo=F}
price_date = INPUT_DATE
date_needed = as.POSIXct(price_date)
date_low = date_needed - days(30)
date_high = date_needed + days(30)

yhat = mean(forecast_lumber[(forecast_lumber$ds < date_high) & (forecast_lumber$ds > date_low), 'yhat'])
yhat_upper = max(forecast_lumber[(forecast_lumber$ds < date_high) & (forecast_lumber$ds > date_low), 'yhat_upper'])
yhat_lower = min(forecast_lumber[(forecast_lumber$ds < date_high) & (forecast_lumber$ds > date_low), 'yhat_lower'])

sr = ((yhat_upper - yhat_lower)/2)/(1.28)
hist_data = rnorm(1000, yhat, sr)

den = density(hist_data)
temp2 = data.frame(matrix(c(den$x, den$y), ncol=2))
colnames(temp2) = c('price', 'frequency')
dygraph(temp2) %>%
  dyOptions(fillGraph = TRUE, fillAlpha = 0.4) %>%
  dyShading(from = yhat-2*sr, to = yhat + 2*sr, color = 'tan')



```


### shingles est for 1000 'shingles'

```{r caption='Shingles price range', render=lemon_print, echo=FALSE}
head(prices_shingle, 2)
```


```{r plot componants and hist shingle, echo=FALSE, warning=FALSE}

prophet_plot_components(m_shingle, forecast_shingle)
dyplot.prophet(m_shingle, forecast_shingle)

```



```{r price_estimate_hist, echo=F}
price_date = INPUT_DATE
date_needed = as.POSIXct(price_date)
date_low = date_needed - days(30)
date_high = date_needed + days(30)

yhat = mean(forecast_shingle[(forecast_shingle$ds < date_high) & (forecast_shingle$ds > date_low), 'yhat'])
yhat_upper = max(forecast_shingle[(forecast_shingle$ds < date_high) & (forecast_shingle$ds > date_low), 'yhat_upper'])
yhat_lower = min(forecast_shingle[(forecast_shingle$ds < date_high) & (forecast_shingle$ds > date_low), 'yhat_lower'])

sr = ((yhat_upper - yhat_lower)/2)/(1.28)
hist_data = rnorm(1000, yhat, sr)

den = density(hist_data)
temp2 = data.frame(matrix(c(den$x, den$y), ncol=2))
colnames(temp2) = c('price', 'frequency')
dygraph(temp2) %>%
  dyOptions(fillGraph = TRUE, fillAlpha = 0.4) %>%
  dyShading(from = yhat-2*sr, to = yhat + 2*sr, color = 'tan')

```


### Total project material cost
```{r total cost, caption='estimate table for full project materials', render=lemon_print, echo=FALSE}
head(prices_shingle+prices_lumber, 2)

```

```{r combined est, echo=FALSE}
total_price = lumber_price

total_price$y = total_price$y*lumber_requirement + shingle_price$y*shingle_reauirement


m_total = prophet(total_price)


future_total = make_future_dataframe(m_total, periods = 365, freq = 'D')

forecast_total = predict(m_total, future_total)

prophet_plot_components(m_total, forecast_total)
dyplot.prophet(m_total, forecast_total)
```

```{r total conf}
price_date = INPUT_DATE
date_needed = as.POSIXct(price_date)
date_low = date_needed - days(30)
date_high = date_needed + days(30)

yhat = mean(forecast_total[(forecast_total$ds < date_high) & (forecast_total$ds > date_low), 'yhat'])
yhat_upper = max(forecast_total[(forecast_total$ds < date_high) & (forecast_total$ds > date_low), 'yhat_upper'])
yhat_lower = min(forecast_total[(forecast_total$ds < date_high) & (forecast_total$ds > date_low), 'yhat_lower'])

sr = ((yhat_upper - yhat_lower)/2)/(1.28)
hist_data = rnorm(1000, yhat, sr)

den = density(hist_data)
temp2 = data.frame(matrix(c(den$x, den$y), ncol=2))
colnames(temp2) = c('price', 'frequency')
dygraph(temp2) %>%
  dyOptions(fillGraph = TRUE, fillAlpha = 0.4) %>%
  dyShading(from = yhat-2*sr, to = yhat + 2*sr, color = 'tan')


```